---
title: "Factor Analysis Project"
author: "S16844"
date: "8/4/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Import Packages
```{r}
library(data.table)
library(factoextra)
library(psych)
library(corrplot)
library(ggplot2)

```

Import data as "df" . I dropped two categorical variables for my analysis
```{r}
df <- fread("bodyfat.csv",header = TRUE)
head(df)
```

Set NA values to zero
```{r}
df[is.na(df)] <- 0
```

Describe the dataset
```{r}
describe(df)
```
Standardized data
```{r}
df_st <- apply(df,2,scale)
df_st
```

KMO Test
```{r}
KMO(df_st)
```
Dimensions of the dataset
```{r}
dim(df_st)
```

Get the covariance matrix of standardized data
```{r}
df_st_cov <- cov(df_st)
df_st_cov
```

Get eigen values and eigen vectors of covariance matrix
```{r}
df_st_cov_eigen <- eigen(df_st_cov)

#eigen values
df_st_cov_eigen$values

#eigen vectors
df_st_cov_eigen$vectors
```
Proportion of variance explained
```{r}
PVE <- df_st_cov_eigen$values / sum(df_st_cov_eigen$values)
PVE
```
Scree Plot
```{r}
scree(df_st)
```

Parallel analysis Scree Plots
```{r}
fa.parallel(df_st,fm="pa",fa="fa")
```
Factor Analysis from 'Principal Component Method'
```{r}
df_st_fa_pc <- fa(df_st_cov ,nfactors = 3,rotate = "none",n.obs = 252 ,covar = TRUE,fm = "pa")
df_st_fa_pc
```
Get unrotated loadings from PC method
```{r}
df_st_fa_pc$loadings
```
Get unrotated PC loadings as dataframe
```{r}
unrotated_pc_loadings <- as.data.frame(unclass(df_st_fa_pc$loadings))
unrotated_pc_loadings
```

Get unrotated PC method communalities as dataframe
```{r}
unrotated_pc_com <- as.data.frame(unclass(df_st_fa_pc$communality))
unrotated_pc_com
```
Factor Analysis from 'Maximum Likelihood Method'
```{r}
df_st_fa_ml <- fa(df_st_cov,nfactors = 3,rotate = "none",n.obs = 252 , covar = TRUE, fm = 'ml')
df_st_fa_ml
```
Get unrotated ML loadings 
```{r}
df_st_fa_ml$loadings
```
Get unrotated ML loadings as dataframe
```{r}
unrotated_ml_loadings <- as.data.frame(unclass(df_st_fa_ml$loadings))
unrotated_ml_loadings
```
Get unrotated Communalities in ML method
```{r}
unrotated_ml_com <- as.data.frame(unclass(df_st_fa_ml$communality))
unrotated_ml_com
```
Rotate the PC method factor loadings using 'Varimax' method
```{r}
library(GPArotation)
df_st_fa_pc_rotate <- fa(df_st_cov ,nfactors = 3,rotate = "varimax",n.obs = 252 ,covar = TRUE,fm = 'pa')
df_st_fa_pc_rotate
```
```{r}
df_st_fa_pc_rotate$loadings
```

Roatate factor loadings as a dataframe from PC method
```{r}
rotated_pc_loadings <- as.data.frame(unclass(df_st_fa_pc_rotate$loadings))
rotated_pc_loadings
```
Get rotated communalities in PC method
```{r}
rotated_pc_com <- as.data.frame(unclass(df_st_fa_pc_rotate$communality))
rotated_pc_com
```












